---
title: "Introduction to FoldSeek"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
library(knitr)
opts_chunk$set(fig.align = "center", 
               out.width = "90%",
               fig.width = 6, fig.height = 5.5,
               dev.args=list(pointsize=10),
               par = TRUE, # needed for setting hook 
               collapse = TRUE, # collapse input & ouput code in chunks
               warning = FALSE)
knit_hooks$set(par = function(before, options, envir)
  { if(before && options$fig.show != "none") 
       par(family = "sans", mar=c(4.1,4.1,1.1,1.1), mgp=c(3,1,0), tcl=-0.5)
})
set.seed(1) # for exact reproducibility
```

## Introduction

`foldSeekR` is an R package developed to simplify protein architecture comparison workflow using AlphaFold and FoldSeek. **This document gives a tour of TestingPackage (version 0.1.0)**. It was written in R Markdown, using the [knitr](https://cran.r-project.org/package=knitr) package for production. See `help(package = "foldSeekR")` for further details. 

To download **foldSeekR**, use the following commands:

``` r
require("devtools")
devtools::install_github("bellabr/foldSeekR", build_vignettes = TRUE)
library("foldSeekR")
```
To list all sample functions available in the package:
``` r
ls("package:foldSeekR")
```

## Getting Started

```{r setup}
library(foldSeekR)
```

Let's take a look at a tyrosine-kinase in mice.

Using the Uniprot accession of interest for this protein, you can pull the AlphaFold prediction.

```{r}
uniprot_accession = "P00520"
prediction <- pull_prediction(uniprot_accession)
prediction$gene
```

```{r}
prediction$organismScientificName
```

```{r}
prediction$uniprotSequence
```

There is a lot of data available here! For our purposes, we only need the URL for visualizing protein architecture.

```{r}
pdb_url <- pull_url(uniprot_accession, "pdb")
pdb_url
```

You can also pull CIF files.

```{r}
cif_url <- pull_url(uniprot_accession, "cif")
cif_url
```

We can visualize this prediction on its own.

```{r}
visualize_prediction(uniprot_accession)
```

This is the structure of the protein of interest. You can move the POV around with your cursor and resize the protein to get a closer look at the folds.

## Integration with FoldSeek

We can submit a job to the FoldSeek search server and wait for the request to be handled.

Due to the inability for vignettes to respond to allow for cache creation, these functions as follows are more difficult than necessary for workflow. The remainder of this tutorial should be performed in your console.

## Console workflow

For this job, the console will prompt you to allow it to create a temporary cache directory.

```r
job <- foldseek("./data/test.cif")
```

Now that we have the job results, we can cache the structure for the top similar hits we want to view.

```r
cache <- cache_top_hits(job, 3, uniprot_accession)
```

Now, with the cache summary, we can access the cached objects to display the most similar protein architectures.

```r
visualize_foldseeks(uniprot_accession, cache)
```

This displays all proteins of interest aligned and together for visualization. This is the final goal of all functions within this package.

Using the powerful visualization tools provided by bio3d and r3dmol, we can visualize all angles of each protein simultaneously. The colour scheme marks similar folds among the proteins.

There is much more work that can be done to build upon this functionality, and much more quality control to be performed.
